The ``encrypt-then-MAC'' transformations

### 2024-07-02 faster-fo-transform/etm-transform/main.tex
Open questions on proving EtM to be plaintext aware

\newcommand{\monospace}{\texttt}
\newcommand{\pke}{\monospace{PKE}}
\newcommand{\keygen}{\monospace{KeyGen}}
\newcommand{\encrypt}{\monospace{E}}
\newcommand{\decrypt}{\monospace{D}}
\newcommand{\kem}{\monospace{KEM}}
\newcommand{\encap}{\monospace{Encap}}
\newcommand{\decap}{\monospace{Decap}}
\newcommand{\etm}{\monospace{EtM}}  % encrypt-then-mac
\newcommand{\mac}{\monospace{MAC}}
\newcommand{\sign}{\monospace{Sign}}
\newcommand{\verify}{\monospace{Verify}}
\newcommand{\pk}{\monospace{pk}}
\newcommand{\sk}{\monospace{sk}}
\newcommand{\pco}{\monospace{PCO}}
\newcommand{\cvo}{\monospace{CVO}}
\newcommand{\lwe}{\monospace{LWE}}
\newcommand{\cbd}{\monospace{CBD}}
\newcommand{\oracle}[1]{\mathcal{O}^{#1}}
\newcommand{\leftsample}{\stackrel{\$}{\leftarrow}}
\newcommand{\llbrack}{[\![}
\newcommand{\rrbrack}{]\!]}
\newcommand{\norm}[1]{\left\lvert #1 \right\rvert}
\newcommand{\round}[1]{\lceil #1 \rfloor}
\newlength{\wdth}
\newcommand{\strike}[1]{\settowidth{\wdth}{#1}\rlap{\rule[.5ex]{\wdth}{.4pt}}#1}
\newcommand{\notrand}{{\strike{\$}}}
\newcommand{\rand}{\$}

@@

\subsection{$\etm$ is plaintext aware}\label{sec:etm-is-pa}

>>>>

\newcommand{\monospace}{\texttt}
\newcommand{\pke}{\monospace{PKE}}
\newcommand{\keygen}{\monospace{KeyGen}}
\newcommand{\encrypt}{\monospace{E}}
\newcommand{\decrypt}{\monospace{D}}
\newcommand{\kem}{\monospace{KEM}}
\newcommand{\encap}{\monospace{Encap}}
\newcommand{\decap}{\monospace{Decap}}
\newcommand{\etm}{\monospace{EtM}}  % encrypt-then-mac
\newcommand{\mac}{\monospace{MAC}}
\newcommand{\sign}{\monospace{Sign}}
\newcommand{\verify}{\monospace{Verify}}
\newcommand{\pk}{\monospace{pk}}
\newcommand{\sk}{\monospace{sk}}
\newcommand{\pco}{\monospace{PCO}}
\newcommand{\cvo}{\monospace{CVO}}
\newcommand{\lwe}{\monospace{LWE}}
\newcommand{\cbd}{\monospace{CBD}}
\newcommand{\oracle}[1]{\mathcal{O}^{#1}}
\newcommand{\leftsample}{\stackrel{\$}{\leftarrow}}
\newcommand{\llbrack}{[\![}
\newcommand{\rrbrack}{]\!]}
\newcommand{\norm}[1]{\left\lvert #1 \right\rvert}
\newcommand{\round}[1]{\lceil #1 \rfloor}
\newlength{\wdth}
\newcommand{\strike}[1]{\settowidth{\wdth}{#1}\rlap{\rule[.5ex]{\wdth}{.4pt}}#1}
\newcommand{\notrand}{{\strike{\$}}}
\newcommand{\negl}[1]{\operatorname{negl}\left(#1\right)}
\newcommand{\rand}{\$}

@@

\subsection{$\etm$ is plaintext aware}\label{sec:etm-is-pa}
I want to prove that $\etm$ is plaintext aware by showing that if $\etm$ is not plaintext aware, then the underlying $\mac$ is broken. For clearer notations we denote unauthenticated ciphertexts by $\sigma \in \mathcal{C}_\pke$ and authenticated ciphertexts by $c = (\sigma, t)$.

If there exists a PA adversary $B$ with output $(\mathcal{L}^{G, H, \mathcal{R}}, C, c=(\sigma, t)) \leftarrow B(\cdot)$ such that for all knowledge extractor $K$

\begin{equation*}
    P \left\lbrack
    K(\mathcal{L}^{G, H, \mathcal{R}}, C, c=(\sigma, t)) \neq \decrypt_\etm(\sk, c)
    \right\rbrack > \negl{\lambda}
\end{equation*}

If $B$ produces a valid ciphertext $c = (\sigma, t)$, then $t$ must be a valid tag of $\sigma$ under $k_\mac = H(m)$ where $m = \decrypt(\sk, \sigma)$.

\begin{enumerate}
    \item How can I express that ``If $B$ knows $m$, then some knowledge extractor can recover $m$''?
    \item If $B$ does not know $m$, then $k_\mac$ is unknown and uniformly random. How can I construct a forgery adversary using $B$?
    \item How do I account for the scenario in which $c$ is invalid?
\end{enumerate}
